---
title: "Chapter_3_data_organization"
author: "Alper"
date: "2023-02-10"
output: html_document
---

# library
```{r message=FALSE, warning=FALSE, include=FALSE}
analysis.path <- getwd()

source("libraries_functions.R") #load the libraries
```

# import additional datasets, demographics, trucount, CMV, immunotype, cytokine
## Import subject information, age, age group, sex
```{r}
## subset the age, age groups and sex data 
vital_subject <- read_excel("/Users/alpercevirgel/Documents/Alper RIVM/1_Complete_analysis_per_project/VITAL database/VITAL database Alper.xlsx",
                            sheet = "subject",
                            skip = 1) %>% 
  mutate_at(c("sex","Age_group_T0"), list(~factor(.))) %>%
  mutate_at(c("subject_identifier"), list(~as.character(.))) %>%
  dplyr::rename(age_group=Age_group_T0,
                age=Age_T0,
                BMI=BMI_T0) %>%
  dplyr::select(subject_identifier,sex,age,age_group) %>%
  as.data.frame()


head(vital_subject) 
```
## Chronic viral infection data
```{r}
## add chronic herpes virus infection data, and create factors for CMV+/- & EBV+/- combinations
chronic_viral_inf <- read_excel("/Users/alpercevirgel/Documents/Alper RIVM/1_Complete_analysis_per_project/VITAL database/VITAL database Alper.xlsx", 
                                       sheet = "CMV_EBV_VZV",
                                       skip = 1,na = "n/a") %>%
  dplyr::select(subject_identifier,MIA_CMV_Titer,MIA_CMV_Seropositivity,MIA_EBV_Titer,MIA_EBV_Seropositivity,MIA_VZV_Titer,MIA_VZV_Seropositivity) %>%
  mutate(CMV_EBV_combined = case_when(
      MIA_CMV_Seropositivity == "1" & MIA_EBV_Seropositivity == "1"  ~ "CMV+EBV+",
      MIA_CMV_Seropositivity == "2" & MIA_EBV_Seropositivity == "2"  ~ "CMV-EBV-",
      MIA_CMV_Seropositivity == "3" | MIA_EBV_Seropositivity == "3"  ~ "Borderline",
      MIA_CMV_Seropositivity == "1" & MIA_EBV_Seropositivity == "2"  ~ "CMV+EBV-",
      MIA_CMV_Seropositivity == "2" & MIA_EBV_Seropositivity == "1"  ~ "CMV-EBV+",
),.after="subject_identifier") %>%
  mutate_at(c("subject_identifier"), list(~as.character(.))) %>%
  mutate_at(c("MIA_CMV_Seropositivity","MIA_EBV_Seropositivity","MIA_VZV_Seropositivity","CMV_EBV_combined"), list(~factor(.))) %>%
  mutate(MIA_CMV_Seropositivity = recode(MIA_CMV_Seropositivity,"3" = "2")) 

head(chronic_viral_inf)
```
## Immunotype
```{r}
immunotype <- readRDS(
  "/Users/alpercevirgel/Documents/Alper RIVM/4_Manuscripts/Chapter_1/R_analysis/vital.truc_influ_d0_cluster.rds") %>%
  dplyr::select(subject_identifier,cluster_number)
```
## immune entropy and stability
```{r}
immune_entropy <- readRDS("~/Documents/Alper RIVM/4_Manuscripts/Chapter_4/R_analysis/immune_entropy_all_VITAL.rds") %>%
  rename(immune_entropy = entropy_cor_T0T5num_med)
immune_stability <- readRDS("~/Documents/Alper RIVM/4_Manuscripts/Chapter_4/R_analysis/immune_stability.rds") %>%
  rename(immune_stability = stability_num)
```

## cytokine
```{r}
cytokine <- read_excel("~/Documents/Alper RIVM/1_Complete_analysis_per_project/VITAL database/VITAL database Alper.xlsx", 
    sheet = "cytokines_final", skip = 1, na="NA") %>%
  mutate_at(c("subject_identifier"), list(~as.character(.))) %>%
  as.data.frame() %>%
  drop_na()
```

# METADATA RADIANT, run only once. This is too create the metadata xlsx to be usd in RADIANT
```{r eval=FALSE, include=FALSE}
## CD4 example, done for CD8 and TCRgd too
read_excel(
  "/Users/alpercevirgel/Documents/Alper RIVM/4_Manuscripts/Chapter_3/R_analysis/RADIANT/CD4/RADIANT Catalyst CD4 metadata.xlsx") %>%
  rename("subject_identifier"="sample_id") %>%
  full_join(vital_subject, by="subject_identifier") %>%
  full_join(immunotype, by="subject_identifier") %>%
  full_join(chronic_viral_inf[c(1,4)], by="subject_identifier") %>%
  full_join(immune_entropy, by="subject_identifier") %>%
  full_join(immune_stability, by="subject_identifier") %>%
  rename("sample_id"="subject_identifier") %>%
  drop_na(file_name) %>%
  write_xlsx("/Users/alpercevirgel/Documents/Alper RIVM/4_Manuscripts/Chapter_3/R_analysis/RADIANT/CD4/RADIANT Catalyst CD4 metadata.xlsx")

```

# flowSOM data, immune subset frequencies import
```{r}
fs_CD4 <- read_excel("/Users/alpercevirgel/Documents/Alper RIVM/4_Manuscripts/Chapter_3/R_analysis/RADIANT/CD4/03 Results/02 FlowSOM_clustering/01 Cluster merging of pqcd/freq_som_merging1.xlsx") %>%
  dplyr::select(sample_id,cluster_id,Freq) %>% 
  pivot_wider(names_from = cluster_id, values_from = Freq) %>%
  rename(subject_identifier=sample_id)
names(fs_CD4)<-gsub(x = names(fs_CD4), pattern = "C", replacement = "CD4c")

fs_CD8 <- read_excel("/Users/alpercevirgel/Documents/Alper RIVM/4_Manuscripts/Chapter_3/R_analysis/RADIANT/CD8/03 Results/02 FlowSOM_clustering/01 Cluster merging of pqcd/freq_som_merging1.xlsx") %>%
  dplyr::select(sample_id,cluster_id,Freq) %>% 
  pivot_wider(names_from = cluster_id, values_from = Freq) %>%
  rename(subject_identifier=sample_id)
names(fs_CD8)<-gsub(x = names(fs_CD8), pattern = "C", replacement = "CD8c")

fs_TCRgd <- read_excel("/Users/alpercevirgel/Documents/Alper RIVM/4_Manuscripts/Chapter_3/R_analysis/RADIANT/TCRgd/03 Results/02 FlowSOM_clustering/01 Cluster merging of pqcd/freq_som_merging1.xlsx") %>%
  dplyr::select(sample_id,cluster_id,Freq) %>% 
  pivot_wider(names_from = cluster_id, values_from = Freq) %>%
  rename(subject_identifier=sample_id)
names(fs_TCRgd)<-gsub(x = names(fs_TCRgd), pattern = "C", replacement = "GDc")


SOMs <- fs_CD4 %>%
  full_join(fs_CD8, by=c("subject_identifier")) %>%
  full_join(fs_TCRgd, by=c("subject_identifier")) 
```

# marker expression median per sample
```{r}
temp_CD4 <- read_excel("/Users/alpercevirgel/Documents/Alper RIVM/4_Manuscripts/Chapter_3/R_analysis/RADIANT/CD4/00 RADIANT files/median_exprss_CD4.xlsx") %>%
  select(-`...1`) %>% 
  as.data.frame() %>%
  mutate(.before=marker ,cell_type=c("CD4"))

temp_CD8 <- read_excel("/Users/alpercevirgel/Documents/Alper RIVM/4_Manuscripts/Chapter_3/R_analysis/RADIANT/CD8/00 RADIANT files/median_exprss_CD8.xlsx") %>%
  select(-`...1`) %>% 
  as.data.frame() %>%
  mutate(.before=marker ,cell_type=c("CD8"))

temp_TCRgd <- read_excel("/Users/alpercevirgel/Documents/Alper RIVM/4_Manuscripts/Chapter_3/R_analysis/RADIANT/TCRgd/00 RADIANT files/median_exprss_TCRgd.xlsx") %>%
  select(-`...1`) %>% 
  as.data.frame() %>%
  mutate(.before=marker ,cell_type=c("TCRgd"))

MFI_celltype <- rbind(temp_CD4,temp_CD8,temp_TCRgd) %>% rename(subject_identifier=sample_id)

exprss_MFI <- rbind(temp_CD4,temp_CD8,temp_TCRgd) %>%
  rename(subject_identifier=sample_id)%>%
  pivot_wider(
              names_from = marker, 
              values_from = median_expression) %>%
  pivot_wider(
    names_from = cell_type,
    values_from = c(CCR4, CCR6, CCR7, CD127, CD159a, CD25, CD27, CD28, CD31, CD38, CD45RA, CD57,
                    CD95, CTLA4, CX3CR1, CXCR3, CXCR5, FOXP3, HLADR, Helios, ICOS, KLGR1, PD1, 
                    TIGIT, panKIR,CD4,CD8),
    names_sep = "_"
  ) %>% 
  select(-c(CD4_CD4,CD4_CD8,CD8_CD8,CD8_CD4))%>%
  as.data.frame()
```

# triple response quartiles
```{r}
trivac_resp_quartiles <- readRDS("/Users/alpercevirgel/Documents/Alper RIVM/4_Manuscripts/Chapter_4/R_analysis/trivac_resp_quartiles.rds") %>%
  mutate(ResponseGroup_Triple2 = recode(ResponseGroup_Triple, "1" = "TQ12", "2" = "TQ12", "3"="TQ34", "4"="TQ34")) %>%
  mutate(ResponseGroup_Triple = recode(ResponseGroup_Triple, "1" = "TQ1", "2" = "TQ2", "3"="TQ3", "4"="TQ4")) %>%
  droplevels()
```

# Manual cell subsets (takes excel from flowjo & organizes for R)
```{r echo=TRUE}
manual_gates <- read_excel("~/Documents/Alper RIVM/1_Complete_analysis_per_project/VITAL Aurora/Analysis/Manual gating/manual_gating2.xls") %>%
  rename("sample_identifier"="...1") %>%
  as.data.frame() %>%
  mutate_all(~str_remove_all(., " %")) %>%
  mutate_all(~str_remove_all(., ".fcs")) %>%
  filter(!str_detect(sample_identifier, "Mean")) %>%
  filter(!str_detect(sample_identifier, "SD")) %>%
  filter(!str_detect(sample_identifier, "control")) %>%
  filter(!str_detect(sample_identifier, "rainbow")) %>%
  filter(!str_detect(sample_identifier, "Bram")) %>%
  mutate_at(vars(-sample_identifier), as.numeric)  %>%
  separate(sample_identifier,
           into = c("batch", "subject_identifier","timepoint"), sep = "_") %>%
  mutate(sample_identifier = ifelse(nchar(subject_identifier) == 2,
                                    paste0("0", subject_identifier), subject_identifier)) %>%
  relocate(where(is.numeric), .after = last_col()) %>%
  mutate(sample_identifier = paste0(sample_identifier, "C")) %>%
  mutate(sample_identifier=ifelse(timepoint == "d0",
                                  paste0("406A", sample_identifier),
                                  paste0("406C", sample_identifier))) %>%
  mutate(timepoint = dplyr::recode(timepoint,
                                   "d0" = "A")) %>%
  dplyr::select(subject_identifier, everything(.)) %>%
  dplyr::select(-c(timepoint,batch,sample_identifier)) 
```

# metas
```{r}
IDs_sub_sam <- manual_gates %>% filter(subject_identifier %!in% c("c1","c2","c3","c4","c5")) %>% select(subject_identifier)


SOMs_meta <- vital_subject %>%
  full_join(immunotype, by="subject_identifier") %>%
  full_join(chronic_viral_inf[c(1,4)], by="subject_identifier") %>%
  full_join(immune_entropy, by="subject_identifier") %>%
  full_join(immune_stability, by="subject_identifier") %>%
  full_join(SOMs, by="subject_identifier")%>%
  as.data.frame() %>%
  left_join(trivac_resp_quartiles) %>%
  drop_na(ResponseGroup_Triple2) %>%
  dplyr::select(where(is.character), where(is.factor), where(is.numeric)) %>%
  filter(subject_identifier %in% IDs_sub_sam$subject_identifier) 

MFI_meta <- exprss_MFI %>%
  full_join(vital_subject) %>%
  full_join(immunotype)%>%
  full_join(chronic_viral_inf[c(1,4)], by="subject_identifier") %>%
  full_join(immune_entropy, by="subject_identifier") %>%
  full_join(immune_stability, by="subject_identifier") %>%
  full_join(trivac_resp_quartiles)  %>%
  drop_na(ResponseGroup_Triple2) %>%
  filter(subject_identifier %in% IDs_sub_sam$subject_identifier)

gating_meta <- vital_subject %>%
  full_join(immunotype, by="subject_identifier") %>%
  full_join(chronic_viral_inf[c(1,4)], by="subject_identifier") %>%
  full_join(immune_entropy, by="subject_identifier") %>%
  full_join(immune_stability, by="subject_identifier") %>%
  full_join(manual_gates, by="subject_identifier")%>%
  as.data.frame() %>%
  dplyr::select(where(is.character), where(is.factor), where(is.numeric)) %>%
  drop_na(subject_identifier) %>%
  left_join(trivac_resp_quartiles) %>%
  drop_na(ResponseGroup_Triple2) %>%
  filter(subject_identifier %in% IDs_sub_sam$subject_identifier)
  

trivac_resp_quartiles <- trivac_resp_quartiles %>% filter(subject_identifier %in% IDs_sub_sam$subject_identifier)
```


